\documentclass[a4paper,11pt]{article}

\author{David Maldonado, $\href{mailto:david.m.maldonado@gmail.com}%
{david.m.maldonado@gmail.com}$}
\title{Boolos and Jeffrey - HW4}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{bussproofs}
\usepackage{cite}
\usepackage[pdftex]{hyperref}
\usepackage{latexsym}
\usepackage{listings}
\usepackage{synttree}
\usepackage{textcomp}
\usepackage{verbatim}
\usepackage{tabu}
\usepackage{tikz}
\usetikzlibrary{trees}
\usetikzlibrary{arrows, automata}
\usepackage[latin1]{inputenc}

\newtheorem{lem}{Lemma}[section]
\newtheorem{thm}{Theorem}[section]
\newtheorem{con}{Conclusion}[section]

\begin{document}

\maketitle

\bigskip


In the following theorems $Q$ stands for \textit{either} quantifier and $Q'$ for its counterpart. $v$ stands for 
a quantified variable, and $F$ and $G$ for first-order logical formulas with no free variables.

% QUESTION 1

\section{some equivalence proofs...}

\begin{thm}$\lnot Qv F \cong Q'v \lnot F $\end{thm}
	
	\begin{proof}
		We'll begin with the first case: 
		
		\begin{equation} \lnot \forall v F \cong \exists v \lnot F \end{equation}
		
		The implication $\lnot \forall v F \implies \exists v \lnot F$ is proven as follows. Let $\mathcal{M}$ be 
		a model that satisfies $\lnot \forall v F$. That means there exists at least one $a \in \mathcal{M}$ 
		that makes $\lnot F$ \textbf{true}, which is precisely the statement on the right-hand side.
		
		The converse implication $\lnot \forall v F \impliedby \exists v \lnot F$ is proven in the same way by
		assuming $\exists v \lnot F$ to be satisfied by $\mathcal{M}$. It follows directly that because there 
		is at least one $a \in \mathcal{M}$ that makes $\lnot F$ \textbf{true} \textit{not all} 
		$a \in \mathcal{M}$  make $F$ \textbf{true} which is the statement on the left-hand side.
		
		\bigskip
		
		For the second case:
		
		\begin{equation} \lnot \exists v F \cong \forall v \lnot F \end{equation}
		
		The implication $\lnot \exists v F \implies \forall v \lnot F$ is proven by first assuming $\lnot \exists v F$ 
		is satisfied by $\mathcal{M}$. With this assumption we can say that there \emph{does not} exist 
		$a \in \mathcal{M}$ such that F is \textbf{true}, this leads to the right-hand statement that for all
		$a \in \mathcal{M}$  $\lnot F$ is \textbf{true}.
		
		The converse implication $\lnot \exists v F \impliedby \forall v \lnot F$ is proven by assuming 
		$\forall v \lnot F$ is satisfied by $\mathcal{M}$. Now we can see that for all $a \in \mathcal{M}$ 
		$\lnot F$ is \textbf{true}, therefore there \emph{does not} exist $a \in \mathcal{M}$ that makes $F$ 
		\textbf{true}, which is the left-hand statement.
	\end{proof}
	
\setcounter{equation}{0}

\pagebreak

\begin{thm}$Qv F \land G \cong Qv(F \land G)$\end{thm}

	\begin{proof}
		For the first case:
	
		\begin{equation} \forall v F \land G \cong \forall v(F \land G) \end{equation}
	
		The implication $\forall v F \land G \implies \forall v(F \land G)$ is proven as follows. 
		Let $\mathcal{M}$ represent a model
		that satisfies $\forall v F \land G$, this means $\mathcal{M}$ satisfies $\forall v F$ 
		and $\mathcal{M}$ satisfies $G$. For $\forall v F$ since all $a \in \mathcal{M}$ make
		$F$ \textbf{true} we can say more generally that $\mathcal{M}$ satisfies $F$. We can now
		combine $F$ and $G$ and say $\mathcal{M}$ satisfies $(F \land G)$. Finally because 
		$v$ does not appear free in $G$ and all $a \in \mathcal{M}$ make $F$ \textbf{true}
		$\mathcal{M}$ satisfies $\forall v(F \land G)$. 
		
		The converse implication $\forall v F \land G \impliedby \forall v(F \land G)$ is proven using 
		the same argument in reverse. If we assume a model $\mathcal{M}$ that satisfies 
		$\forall v(F \land G)$ we can say that all $a \in \mathcal{M}$ make $F \land G$ \textbf{true}.
		We can then say that $\mathcal{M}$ satisfies $F$ and $\mathcal{M}$ satisfies $G$. 
		(kind of got stuck here, so we have $F$ and $G$ seperately satisfied by $\mathcal{M}$, 
		how do I recombine them on the other side to make $\forall v Fv \land G$ ????)
		
		\begin{equation} \exists v F \land G \cong \exists v(F \land G) \end{equation}
		
		In the second case we'll start with the implication $\exists vF \land G \implies \exists v(F \land G)$.
		We'll assume a model $\mathcal{M}$ that satisfies $\exists vF \land G$. This means that 
		$\mathcal{M}$ satisfies $\exists vF$ and $\mathcal{M}$ satisfies $G$. (I'm assuming I'm going to 
		use the same argument as $\forall$ but if I can only say there exists an $a \in \mathcal{M}$ 
		that makes $F$ \textbf{true} can I generalize and say $\mathcal{M}$ satisfies F? or do I have to think 		about another strategy?)
		
 	\end{proof}
	
\setcounter{equation}{0}

% QUESTION 2	
	
\section{proof of prenex normal form}

\begin{thm} Where \textbf{prenex normal form} is a logical formula where all the quantifiers are written as a string at the front and range over the quantifier-free matrix, every formula in first-order logic has an equivalent prenex normal form. \end{thm}

	
	\begin{proof}
		We will proceed by induction on the complexity of the formula. 
		Let us first agree on the following equivalences:
		
		\begin{equation} \lnot Qv F \cong Q'v \lnot F \end{equation}
		\begin{equation} Qv F \land G \cong Qv(F \land G) \end{equation}
		\begin{equation} G \land Qv F \cong Qv(G \land F) \end{equation}	
		\begin{equation} Qv F \lor G \cong Qv(F \lor G) \end{equation}
		\begin{equation} G \lor Qv F \cong Qv(G \lor F) \end{equation}
		\begin{equation} Qv F  \rightarrow G \cong Q'v(F \rightarrow G) \end{equation}
		\begin{equation} G \rightarrow Qv F \cong Qv(G \rightarrow F) \end{equation}
		
		\bigskip
		
		and the \textbf{principle of substitution of equivalents}:
		
		\begin{equation} Qv F \cong Qw F_{v} w. \end{equation}
		
		\bigskip
		
		Our \textit{base case} is the logical formula with no quantified variables. This formula
		is automatically prenex normal form. The \textit{inductive step} assumes that every formula $F$
		with $n$ or fewer logical symbols is equivalent to a formula $F'$ in prenex normal form. 
		Now we consider a formula with $n + 1$ logical symbols. This well-formed formula is
		one of the following:
		
		\begin{equation*} \lnot \phi \end{equation*}
		\begin{equation*} \phi \land \psi \end{equation*}
		\begin{equation*} \phi \lor \psi \end{equation*}
		\begin {equation*} \phi \rightarrow \psi \end{equation*}
		
		\smallskip
		
		By repeated applications of (1) - (7) and the usage of (8) to change bound variables 
		it is possible to convert any of these formulas into prenex normal form as we'll show below.
		
		\textit{Case 1.} For $\lnot \phi$ if $\phi$ has no quantifiers it is in prenex normal form. 
		If $\phi$ has one or more quantifiers by the induction hypothesis it is in prenex normal form 
		($Q_{1}v \dots Q_{n}v F$). The negation can be moved to the matrix by $n$ applications 
		of (1) using (8) to change bound variables when needed.
		
		\textit{Case 2.} For $\phi \land \psi$ if $\phi$ and $\psi$ both have no quantifiers $\phi \land \psi$
		is in prenex normal form. If either (or both) have quantifiers mixed applications of (2) and (3), 
		depending on which side the quantifier is on, and use of (8) to change bound variables when needed 
		will yield a formula in prenex normal form.
		
		\textit{Case 3.} For $\phi \lor \psi$ the proof is the same as \textit{Case 2} except using (4) and (5)
		as the main operations.
		
		\textit{Case 4.} For $\phi \rightarrow \psi$ the proof is the same as \textit{Case 2} and \textit{Case 3}
		except using (6) and (7) as the main operations.
		
		
	\end{proof}
	
\end{document}